Toronto Island Ferry Tickets: Sales, Redemption Analysis, and Strategic Solutions

### 2024-09-27 paper/paper.qmd
update README, paper, references

---
title: "Toronto Island Ferry Tickets: Sales, Redemption Analysis, and Strategic Solutions"
subtitle: ""
author: 
  - Yun Chu
thanks: "code and data supporting this analaysis can be found here: https://github.com/chuyun2024/Analyzing-Toronto-Islands-Ferry-Tickets-Sales-and-Redemption-Count-Data"
date: 26 September 2024
date-format: long
abstract: "This paper presents a comprehensive analysis of the Toronto Islands ferry ticket sales and redemption dataset. The analysis begins with an overview of the dataset, followed by an examination of the cleaned data. In the results section, we explore the relationship between sales and redemption data, the distribution of redemption counts, and the monthly trends in redemption activity. Based on the findings, we offer recommendations to optimize ferry operations, focusing on pricing strategies, staffing allocations, and departure schedules. The paper concludes by discussing the limitations of the analysis and identifying potential areas for future improvement."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Ferries to Toronto Island Park operate year-round, transporting passengers between the Jack Layton Ferry Terminal and Center Island, Hanlan's Point, and Ward's Island, or in the reverse direction from the islands to the terminal. Currently, the ferries operate on four different schedules, each corresponding to one of the four seasons of the year [@toronto_ferry_website].

These islands are popular destinations for both tourists and residents of Toronto. By analyzing the sales and redemption count data of the ferry service, we can gain insights into visitor trends, helping to optimize resource allocation and revenue management.Identifying peak ferry usage seasons enables a better understanding of when the islands are most popular. This information can be instrumental for governments in resource allocation and operational planning, such as adjusting ferry schedules and staffing levels throughout the year. Additionally, analyzing sales data over time helps uncover revenue trends, providing valuable insights into the financial health of the ferry service and informing pricing strategies.

@@

The remainder of this paper is structured as follows. @sec-data discusses the features of the raw and cleaned data, along with a summary of the cleaned data. @sec-result investigates the relationship between sales and redemption data using graphs, linear regression, and correlation analysis. It also explores the distribution of redemption counts by frequency and counts. @sec-discussion further elaborates on the implications and applications of the findings from @sec-result, addresses the limitations of the analysis, and suggests potential improvements.

@@

## Raw Data & Measurement

The data used in this paper is sourced from the Toronto Open Data portal and was imported using the opendatatoronto library (Gelfand, 2022). The raw dataset contains sales and redemption count data, recorded at 15-minute intervals, collected from both online and POS-kiosk sales. The dataset contains four variables: id, timestamp, sales count and redemption count. The id variable is a sequential number that records the order of the data entries. The dataset spans from July 11, 2023, to September 20, 2024, at the time of download [@toronto_ferry_tickets]. Similar datasets were not found online.

@@

- tidyverse [@tidyverse]
- lubridate [@lubridate]
- dplyr [@dplyr]
- knitr [@knitr]
- scales [@scales]

## Cleaned Data

@@

```{r, fig.width = 8, fig.height = 5}
#| label: fig-Sales_and_Redemption_vs_Date
#| fig-cap: line plot of sales and redemption data with respect to time
#| echo: false

@@

## Summary Statistic

@@

```{r}
#| label: tbl-mean_sales_and_redemption
#| echo: false
#| warning: false
#| message: false
# Calculate summary statistics
summary_statistic <- data |>
  summarise(
    Min_Sales = min(daily_sales_count),
    Max_Sales = max(daily_sales_count),
    Mean_Sales = mean(daily_sales_count),
    Std_Sales = sd(daily_sales_count),
    Min_Redemption = min(daily_redemption_count),
    Max_Redemption = max(daily_redemption_count),
    Mean_Redemption = mean(daily_redemption_count),
    Std_Redemption = sd(daily_redemption_count)
  )

@@

# Display the summary table with row and column headings
kable(summary_table,
  format = "latex", row.names = TRUE, col.names = c("Min", "Max", "Mean", "Standard Deviation"),
  caption = "Summary Statistics for Sales and Redemption Count",
  align = "ccccc"
) |>
  kable_styling(full_width = FALSE, position = "left")
```

The summary statistics for sales and redemption counts, presented in Table @tbl-mean_sales_and_redemption, reveal notable similarities across all metrics.

@@

# Result{#sec-result}

## Investigating the Relationship between the Sales and Redemption Data{#sec-result1}

@@

# Create a scatter plot using ggplot2 to visualize the relationship between sales and redemption
ggplot(data, aes(x = daily_sales_count, y = daily_redemption_count)) + # Set the data and aesthetic mappings; daily sales on the x-axis and daily redemption on the y-axis
  geom_point(alpha = 0.5, color = "#008080") + # Add points to the plot to represent each observation
  labs(
    title = "Scatter Plot of Sales vs Redemption", # Add a title to the plot
    x = "Ticket Sales Count", # Label for the x-axis
    y = "Ticket Redemption Count"
  ) + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5))
```

A linear relationship can be seen from the scatter plot. To further investigate this relationship, we performed a linear regression with redemption count as the response variable and sales count as the predictor variable. As shown in Table @tbl-linear_regression_and_correlation_result, the R-squared statistic is 0.9882, and the correlation coefficient between redemption count and sales count is 0.994. These values indicate a strong linear correlation between the two variables.

@@

## Investigating the Frequency of Redemption Count Data

@fig-Daily_Redemption_Histogram illustrates that for most days, the redemption count remains below 10,000. However, there are some days where the redemption count peaks at 20,234.

For our resource allocation needs, utilizing monthly data would be sufficient to create an effective staffing schedule. Daily scheduling may not be as feasible or efficient, making a monthly approach more practical for our purposes. The monthly redemption distribution is analysed @sec-monred.

@@

@@

From July to September 2023 and from May to September 2024, the monthly redemption counts exceed 100,000. In contrast, from October 2023 to April 2024, the redemption counts fall below this threshold each month.

Given these observations, it would be prudent to design two distinct staffing schedules for the ferries: one for peak periods when redemption counts are high and another for off-peak periods when counts are lower. This approach enables more efficient staff allocation based on expected demand.

## Investigating the Monthly Redemption Data{sec-monred}

@@

# Discussion{#sec-discussion}

## Impact of Pricing Strategy on Revenue Optimization {#sec-first-point}

@@

## Efficient Staffing Solutions for Peak and Off-Peak Seasons

@@

## Evaluation of Staffing and Departure Schedules in Relation to Visitor Demand

>>>>

---
title: "Toronto Island Ferry Tickets: Sales, Redemption Analysis, and Strategic Solutions"
subtitle: ""
author: 
  - Yun Chu
thanks: "code and data supporting this analaysis can be found here: https://github.com/chuyun2024/Analyzing-Toronto-Islands-Ferry-Tickets-Sales-and-Redemption-Count-Data"
date: 26 September 2024
date-format: long
abstract: "This paper presents a comprehensive analysis of the Toronto Islands ferry ticket sales and redemption dataset. The analysis begins with an overview of the dataset, followed by an examination of the cleaned data. In the results section, the relationship between sales and redemption data, the distribution of redemption counts, and the monthly trends in redemption activity are explored. Based on the findings, recommendations to optimize ferry operations, focusing on pricing strategies, staffing allocations, and departure schedules are offered. The paper concludes by discussing the limitations of the analysis and identifying potential areas for future improvement."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Ferries to Toronto Island Park operate year-round, transporting passengers between the Jack Layton Ferry Terminal and Center Island, Hanlan's Point, and Ward's Island, or in the reverse direction from the islands to the terminal. Round-trip tickets are priced at 9.11 Canadian Dollars for adults, with discounted rates available for seniors, youth, juniors, and infants. Currently, the ferries operate on four different schedules, each corresponding to one of the four seasons of the year [@toronto_ferry_website].

These islands are popular destinations for both tourists and residents of Toronto. By analyzing the sales and redemption count data of the ferry service, we can gain insights into visitor trends, helping to optimize resource allocation and revenue management. Identifying peak ferry usage seasons enables a better understanding of when the islands are most popular. This information can be instrumental for governments in resource allocation and operational planning, such as adjusting ferry schedules and staffing levels throughout the year. Additionally, analyzing sales data over time helps uncover revenue trends, providing valuable insights into the financial health of the ferry service and informing pricing strategies.

@@

The remainder of this paper is structured as follows. @sec-data1 - @sec-data3 discusses the features of the raw and cleaned data, along with a summary of the cleaned data. @sec-result1 - @sec-result3 investigates the relationship between sales and redemption data using graphs, linear regression, and correlation analysis. It also explores the distribution of redemption counts by frequency and counts. @sec-dis1 - @sec-dis3 further elaborates on the implications and applications of the findings from @sec-result1 - @sec-result3, addresses the limitations of the analysis, and suggests potential improvements.

@@

## Raw Data & Measurement{#sec-data1}

The data used in this paper is sourced from the Toronto Open Data portal and was imported using the opendatatoronto library (Gelfand, 2022). The raw dataset contains sales and redemption count data, recorded at 15-minute intervals, collected from both online and POS-kiosk sales. The dataset comprises four variables: ID, timestamp, sales count, and redemption count. The ID variable is a sequential number that records the order of the data entries. Both sales and redemption counts are measured per person, with the redemption count indicating how many individuals successfully redeemed their tickets and boarded the ferries. The sales count represents the total number of tickets sold at kiosks and online. The dataset spans from July 11, 2023, to September 20, 2024, at the time of download [@toronto_ferry_tickets]. Similar datasets were not found online.

@@

- tidyverse [@tidyverse]
- lubridate [@lubridate]
- dplyr [@dplyr]
- knitr [@knitr]
- scales [@scales]
- kableExtra [@kableExtra]

## Cleaned Data{#sec-data2}

The data has been cleaned by removing the hour and minute portions from the Timestamp variable and eliminating the ID variable. Additionally, the sales and redemption count data have been recalculated based on the date rather than the previous 15-minute time intervals.

@@

```{r, fig.width = 7, fig.height = 4}
#| label: fig-Sales_and_Redemption_vs_Date
#| fig-cap: line plot of sales and redemption data with respect to time
#| echo: false

@@

## Summary Statistic{#sec-data3}

@@

```{r}
#| label: tbl-mean_sales_and_redemption
#| echo: false
#| warning: false
#| message: false
#| fig.show: hold

# Calculate summary statistics
summary_statistic <- data |>
  summarise(
    Min_Sales = min(daily_sales_count),
    Max_Sales = max(daily_sales_count),
    Mean_Sales = mean(daily_sales_count),
    Std_Sales = sd(daily_sales_count),
    Min_Redemption = min(daily_redemption_count),
    Max_Redemption = max(daily_redemption_count),
    Mean_Redemption = mean(daily_redemption_count),
    Std_Redemption = sd(daily_redemption_count)
  )

@@

# Display the summary table with row and column headings
kable(summary_table,
  format = "latex",
  row.names = TRUE,
  col.names = c("Min", "Max", "Mean", "Standard Deviation"),
  caption = "Summary Statistics for Sales and Redemption Count",
  align = "ccccc"
) |>
  kable_styling(full_width = FALSE, position = "left")

# Note about the table: Setting the format = "latex" resulting that the table does not show under the code chunk, but when rendering, the table shows up fine. If you want to see the table under the code, change format = "html" and it will show up.
```

The summary statistics for sales and redemption counts, presented in @tbl-mean_sales_and_redemption, reveal notable similarities across all metrics.

@@

## Analyais - Investigating the Relationship between the Sales and Redemption Data{#sec-result1}

@@

# Create a scatter plot using ggplot2 to visualize the relationship between sales and redemption
ggplot(data, aes(x = daily_sales_count, y = daily_redemption_count)) + # Set the data and aesthetic mappings; daily sales on the x-axis and daily redemption on the y-axis
  geom_point(alpha = 0.5, color = "#008080") + # Add points to the plot to represent each observation
  labs(
    title = "Scatter Plot of Sales vs Redemption of Toronto's Ferries", # Add a title to the plot
    x = "Ticket Sales Count", # Label for the x-axis
    y = "Ticket Redemption Count"
  ) + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5))
```

A linear relationship can be seen from the scatter plot. To further investigate this relationship, a linear regression with redemption count as the response variable and sales count as the predictor variable is performed. As shown in @tbl-linear_regression_and_correlation_result, the R-squared statistic is 0.9882, and the correlation coefficient between redemption count and sales count is 0.9941. These values indicate a strong linear correlation between the two variables.

@@

## Analysis - Examining the Frequency of Redemption Count Data{#sec-result2}

@fig-Daily_Redemption_Histogram illustrates that for most days, the redemption count remains below 10,000. The 95 percentile of the redemption count is 16075.4. There are some days where the redemption count peaks at 20,234.

For our resource allocation needs, utilizing monthly data would be sufficient to create an effective staffing schedule. Daily scheduling may not be as feasible or efficient, making a monthly approach more practical for our purposes. The monthly redemption distribution is analysed @sec-result3.

@@

## Analysis - Exploring the Monthly Redemption Data{#sec-result3}

@@

From July to September 2023 and from May to September 2024, the monthly redemption counts exceed 100,000. In contrast, from October 2023 to April 2024, the redemption counts fall below this threshold.

Given these observations, it would be prudent to design two distinct staffing and departure schedules for the ferries: one for peak periods when redemption counts are high and another for off-peak periods when counts are lower. This approach enables more efficient staff and ferries allocation and based on expected demand.

@@

## Discussion - Impact of Pricing Strategy on Revenue Optimization {#sec-dis1}

@@

## Discussion - Efficient Staffing Solutions for Peak and Off-Peak Seasons{#sec-dis2}

@@

## Discussion - Evaluation of Staffing and Departure Schedules in Relation to Visitor Demand{#sec-dis3}
